But seriously: what do algorithms want?

### 2023-02-06 ART_Soloducha/Soloducha.tex
Final commit before sending to the editors for corrections

\begin{artengenv}{Krzysztof Sołoducha}
	{Analysis of the implications of the Moral Machine project as an implementation of the concept of coherent extrapolated volition for building clustered trust in autonomous machines}
	{Analysis of the implications of the Moral Machine project\ldots}
	{Analysis of the implications of the Moral Machine project as an implementation of the concept of coherent extrapolated volition for building clustered trust in autonomous machines}
	{Military University of Technology}
	{In this paper, we focus on the analysis of Eliezer Yudkowsky's concept of ``coherent extrapolated volition'' (CEV) as a~response to the need for a~post-conventional, persuasive morality that meets the criteria of active trust in the sense of Anthony Giddens, which could be used in the case of autonomous machines. Based on the analysis of the results of the Moral Machine project, we formulate some guidelines for transformation of the idea of a~coherent extrapolated volition into the concept of a~coherent, extrapolated and clustered volition. The argumentation used in the paper is intended to show that the idea of CEV transformed into its clustered version can be used to build a~technically and socially efficient decision-making pattern database for autonomous machines.
	}
	{ethics of artificial intelligence, ethics of autonomous machines, trust in artificial intelligence, moral machine project, coherent extrapolated volition.}

@@

The problem of the ethics of autonomous machines is a~philosophical issue that has emerged together with the dangers of the development of modern technologies that allow the autonomisation of machine operation
%\label{ref:RNDCNwvoNhD5U}(Pasquale, 2016).
\parencite[][]{pasquale_black_2016}. %
 Of course, the classic of considerations on this issue is Isaac Asimov 
%\label{ref:RND8EDGzYgvVE}(2004)
\parencite*[][]{asimov_runaround_2004} %
 and his reflections on the ethics of robots, but at the moment the discussion on this topic is determined primarily by the dynamic development of unsupervised machine learning methods 
%\label{ref:RNDZsSSj9nxYT}(Gryz, 2021).
\parencite[][]{gryz_sztuczna_2021}. %
 They have gained a~second wind (having been theoretically developed in the 1980s) thanks to the development of the Internet and access to large sets of data on which machines can learn and improve algorithms on their own, using formal rules of statistical reasoning and without any kind of supervision 
%\label{ref:RND2WtqnrZTv3}(Burrell, 2016).
\parencite[][]{burrell_how_2016}.%

>>>>

\begin{artengenv}{Krzysztof Sołoducha}
	{Analysis of the implications of the Moral Machine project as an implementation of the concept of coherent extrapolated volition for building clustered trust in autonomous machines}
	{Analysis of the implications of the Moral Machine project\ldots}
	{Analysis of the implications of the Moral Machine project as\\an implementation of the concept of coherent extrapolated volition\\for building clustered trust in autonomous machines}
	{Military University of Technology}
	{In this paper, we focus on the analysis of Eliezer Yudkowsky's concept of ``coherent extrapolated volition'' (CEV) as a~response to the need for a~post-conventional, persuasive morality that meets the criteria of active trust in the sense of Anthony Giddens, which could be used in the case of autonomous machines. Based on the analysis of the results of the Moral Machine project, we formulate some guidelines for transformation of the idea of a~coherent extrapolated volition into the concept of a~coherent, extrapolated and clustered volition. The argumentation used in the paper is intended to show that the idea of CEV transformed into its clustered version can be used to build a~technically and socially efficient decision-making pattern database for autonomous machines.
	}
	{ethics of artificial intelligence, ethics of autonomous machines, trust in artificial intelligence, moral machine project, coherent extrapolated volition.}

@@

\lettrine[loversize=0.13,lines=2,lraise=-0.03,nindent=0em,findent=0.2pt]%
{T}{}he problem of the ethics of autonomous machines is a~philosophical issue that has emerged together with the dangers of the development of modern technologies that allow the autonomisation of machine operation
%\label{ref:RNDCNwvoNhD5U}(Pasquale, 2016).
\parencite[][]{pasquale_black_2016}. %
 Of course, the classic of considerations on this issue is Isaac Asimov 
%\label{ref:RND8EDGzYgvVE}(2004)
\parencite*[][]{asimov_runaround_2004} %
 and his reflections on the ethics of robots, but at the moment the discussion on this topic is determined primarily by the dynamic development of unsupervised machine learning methods 
%\label{ref:RNDZsSSj9nxYT}(Gryz, 2021).
\parencite[][]{gryz_sztuczna_2021}. %
 They have gained a~second wind (having been theoretically developed in the 1980s) thanks to the development of the Internet and access to large sets of data on which machines can learn and improve algorithms on their own, using formal rules of statistical reasoning and without any kind of supervision 
%\label{ref:RND2WtqnrZTv3}(Burrell, 2016).
\parencite[][]{burrell_how_2016}.%

### 2023-02-06 ART_Toscano/Toscano.tex
Noh article added
Final commit before sending to the editors for corrections

\section*{Initial Definitional Attempts}
Oversimplified definitions of an algorithm are currently available and frequently used, but an algorithm is neither a~recipe nor a~rigidly constrained and procedural formulation. Limited conceptions that represent them as a~sort of entity or thing, a~series of steps that need to be applied, or a~simple technique that homogenizes a~process, lead to weak understandings of the deeper processes, transactions and dynamics that are at stake. Indeed, an algorithm can be a~problem-solving device, and this feature in itself can become a~point of entry to a~more complex analysis. After all, for engineers and computer scientists, ``an algorithm is an abstract, formalized description of a~computational procedure''
%\label{ref:RNDhginqouzXT}(Dourish, 2016, p.3).
\parencite[][p.3]{dourish_algorithms_2016}. %
 But even if sleek and apparently elegant, the problem with this definitional reduction is twofold.

@@

%\section*{References}
%Adams, F. and Aizawa, K., 2008. \textit{The Bounds of Cognition}. 1st ed. Malden, Ma: Blackwell Publishing.
%
%Adams, F. and Aizawa, K., 2010. Defending the Bounds of Cognition. In: R. Menary, ed. \textit{The Extended Mind}. [online] The MIT Press. pp.67–80. https://doi.org/10.7551/mitpress/9780262014038.003.0004.
%
%Aiello, L.C. and Wheeler, P., 1995. The Expensive-Tissue Hypothesis: The Brain and the Digestive System in Human and Primate Evolution. \textit{Current Anthropology}, [online] 36(2), pp.199–221. https://doi.org/10.1086/204350.
%
%Bijker, W.E., 1999. \textit{Of Bicycles, Bakelites, and Bulbs: Toward a~Theory of Sociotechnical Change}. 3rd ed. Inside technology. Cambridge, MA: MIT Press.
%
%Blumer, H., 1986. \textit{Symbolic Interactionism: Perspective and Method}. Berkeley: University of California Press.
%
%Bourdieu, P., 1972. \textit{Outline of a~Theory of Practice}. Cambridge, MA: Cambridge University Press.
%
%Brandom, R.B., 1994. \textit{Making It Explicit: Reasoning, Representing, and Discursive Commitment}. Cambridge, MA: Harvard University Press.
%
%Brandom, R.B., 2000. \textit{Articulating Reasons: An Introduction to Inferentialism}. Cambridge, MA: Harvard University Press.
%
%Brentano, F., 1995. \textit{Psychology from an Empirical Standpoint [1874}]. International Library of Philosophy. Translated by A.C. Rancurello, D.B. Terrell and L.L. McAlister London; New York: Routledge.
%
%Bryant, L.R., 2010. Onticology–A Manifesto for Object-Oriented Ontology Part I. \textit{Larval Subjects}. Available at: {\textless}https://larvalsubjects.wordpress.com/2010/01/12/object-oriented-ontology-a-manifesto-part-i/{\textgreater} [Accessed 18 January 2023].
%
%Callon, M., 1986. Some Elements of a~Sociology of Translation: Domestication of the Scallops and the Fishermen of Saint Brieuc Bay. In: J. Law, ed. \textit{Power, action, and belief: a~new sociology of knowledge}?, Sociological review monograph. London; Boston: Routledge \& Kegan Paul. pp.196–233.
%
%Cash, M., 2008. Thoughts and oughts. \textit{Philosophical Explorations}, [online] 11(2), pp.93–119. https://doi.org/10.1080/13869790802015635.
%
%Cash, M., 2009. Normativity is the mother of intention: Wittgenstein, normative practices and neurological representations. \textit{New Ideas in Psychology}, [online] 27(2), pp.133–147. https://doi.org/10.1016/j.newideapsych.2008.04.010.
%
%Certeau, M. de, 1984. \textit{The Practice of Everyday Life}. Translated by S. Rendall Berkeley \& Los Angeles, CA: University of California Press.
%
%Chant, S.R., Hindriks, F. and Preyer, and G. eds., 2014. \textit{From Individual to Collective Intentionality: New Essays}. Oxford, New York: Oxford University Press.
%
%Clark, A., 1997. \textit{Being There: Putting Brain, Body, and World Together Again}. Cambridge; London: The MIT Press.
%
%Clark, A., 2003. \textit{Natural-Born Cyborgs: Minds, Technologies, and the Future of Human Intelligence}. Oxford; New York: Oxford University Press.
%
%Clark, A., 2015. What ‘Extended Me' knows. \textit{Synthese}, [online] 192(11), pp.3757–3775. https://doi.org/10.1007/s11229-015-0719-z.
%
%Clark, A. and Chalmers, D., 1998. The extended mind. \textit{Analysis}, [online] 58(1), pp.7–19. Available at: {\textless}https://www.jstor.org/stable/3328150{\textgreater} [Accessed 3 June 2019].
%
%Clemens, R. ed., 2016. \textit{The Voynich Manuscript}. New Haven; London: Yale University Press.
%
%Clowes, R.W., 2019. Screen reading and the creation of new cognitive ecologies. \textit{AI \& Society}, [online] 34(4), pp.705–720. https://doi.org/10.1007/s00146-017-0785-5.
%
%Cole, M. and Engeström, Y., 1993. A~cultural-historical approach to distributed cognition. In: G. Salomon, ed. \textit{Distributed Cognitions: Psychological and Educational Considerations}, Learning in doing, 1st ed. Cambridge; New York: Cambridge University Press. pp.1–46.
%
%Connerton, P., 1989. \textit{How Societies Remember}. 1st ed. [online] Cambridge University Press. https://doi.org/10.1017/CBO9780511628061.
%
%Cooren, F., Taylor, J.R. and Van Every, E.J., 2006. \textit{Communication as Organizing: Empirical and Theoretical Approaches into the Dynamic of Text and Conversation}. LEA's communication series. Mahwah, NJ: Lawrence Erlbaum.
%
%Dennett, D.C., 1991. \textit{Consciousness Explained}. Boston: Little, Brown and Company.
%
%Douglas, M., 1986. \textit{How Institutions Think}. 1st ed. The Frank W. Abrams lectures. Syracuse, NY: Syracuse University Press.
%
%Dourish, P., 2016. Algorithms and their others: Algorithmic culture in context. \textit{Big Data \& Society}, [online] 3(2), pp.1–11. https://doi.org/10.1177/2053951716665128.
%
%Dunbar, R.I.M., 1993. Coevolution of neocortical size, group size and language in humans. \textit{Behavioral and Brain Sciences}, [online] 16(4), pp.681–694. https://doi.org/10.1017/S0140525X00032325.
%
%Elster, J., 1983. \textit{Explaining Technical Change: A~Case Study in the Philosophy of Science}. Studies in Rationality and Social Change. [online] Cambridge[202F?]; London et al.: Cambridge University Press. Available at: {\textless}http://archive.org/details/Elster1983ExplainingTechnicalChangeACaseStudyInThePhilosophyOfScienceBook{\textgreater} [Accessed 18 January 2023].
%
%Finn, E., 2017. \textit{What Algorithms Want: Imagination in the Age of Computing}. 1st ed. Cambridge, MA: The MIT Press. https://doi.org/10.7551/mitpress/10766.001.0001.
%
%Gallagher, S., 2005. \textit{How the Body Shapes the Mind}. Oxford; New York: Clarendon Press.
%
%Gallagher, S., 2013. The socially extended mind. \textit{Cognitive Systems Research}, [online] 25–26, pp.4–12. https://doi.org/10.1016/j.cogsys.2013.03.008.
%
%Gallagher, S. and Miyahara, K., 2012. Neo-Pragmatism and Enactive Intentionality. In: J. Schulkin, ed. \textit{Action, Perception and the Brain: Adaptation and Cephalic Expression}, New Directions in Philosophy and Cognitive Science. [online] London: Palgrave Macmillan UK. pp.117–146. https://doi.org/10.1057/9780230360792\_6.
%
%Geertz, C., 1973. \textit{The Interpretation of Cultures: Selected Essays}. New York: Basic Books.
%
%Glock, H.-J., 1986. Vygotsky and Mead on the Self, Meaning and Internalisation. \textit{Studies in Soviet Thought}, [online] 31(2), pp.131–148. Available at: {\textless}https://www.jstor.org/stable/20100086{\textgreater} [Accessed 18 January 2023].
%
%Harman, G., 2002. \textit{Tool-Being: Heidegger and the Metaphysics of Objects}. 1st ed. Chicago: Open Court.
%
%Hollan, J., Hutchins, E. and Kirsh, D., 2000. Distributed cognition: toward a~new foundation for human-computer interaction research. \textit{ACM Transactions on Computer-Human Interaction}, [online] 7(2), pp.174–196. https://doi.org/10.1145/353485.353487.
%
%Hutchins, E., 1995. How a~cockpit remembers its speeds. \textit{Cognitive Science}, [online] 19(3), pp.265–288. https://doi.org/10.1207/s15516709cog1903\_1.
%
%Hutchins, E., 2010. Cognitive ecology. \textit{Topics in Cognitive Science}, 2(4), pp.705–715. https://doi.org/10.1111/j.1756-8765.2010.01089.x.
%
%Jones, C.G., Lawton, J.H. and Shachak, M., 1997. Positive and Negative Effects of Organisms as Physical Ecosystem Engineers. \textit{Ecology}, [online] 78(7), pp.1946–1957. https://doi.org/10.1890/0012-9658(1997)078[1946:PANEOO]2.0.CO;2.
%
%Knorr Cetina, K., 1997. Sociality with Objects: Social Relations in Postsocial Knowledge Societies. \textit{Theory, Culture \& Society}, [online] 14(4), pp.1–30. https://doi.org/10.1177/026327697014004001.
%
%Laland, K.N., Odling-Smee, J. and Feldman, M.W., 2000. Niche construction, biological evolution, and cultural change. \textit{Behavioral and Brain Sciences}, [online] 23(1), pp.131–146. https://doi.org/10.1017/S0140525X00002417.
%
%Latour, B., 1992. Where Are the Missing Masses? The Sociology of a~Few Mundane Artifacts. In: W.E. Bijker and J. Law, eds. \textit{Shaping technology/building society: studies in sociotechnical change}, Inside technology. Cambridge, MA: MIT Press. pp.225–258.
%
%Latour, B., 2005. \textit{Reassembling the Social: An Introduction to Actor-Network-Theory}. Clarendon lectures in management studies. Oxford; New York: Oxford University Press.
%
%Lave, J., 1988. \textit{Cognition in Practice: Mind, Mathematics and Culture in Everyday Life}. [online] Cambridge: Cambridge University Press. https://doi.org/10.1017/CBO9780511609268.
%
%Law, J., 2004. \textit{After Method: Mess in Social Science Research}. International library of sociology. London; New York: Routledge.
%
%Loh, K.K. and Kanai, R., 2016. How Has the Internet Reshaped Human Cognition? \textit{The Neuroscientist}, [online] 22(5), pp.506–520. https://doi.org/10.1177/1073858415595005.
%
%Malafouris, L., 2005. The Cognitive Basis of Material Engagement: Where Brain, Body and Culture Conflate. In: E. DeMarrais, C. Gosden and C. Renfrew, eds. \textit{Rethinking Materiality: Engagement of Mind with Material World}. Cambridge: McDonald Institute for Archaeological Research. pp.53–61.
%
%Malafouris, L., 2008. At the Potter's Wheel: An Argument for Material Agency. In: C. Knappett and L. Malafouris, eds. \textit{Material Agency: Towards a~Non-Anthropocentric Approach}. [online] Berlin: Springer. pp.19–36. https://doi.org/10.1007/978-0-387-74711-8\_2.
%
%Malafouris, L., 2010. Metaplasticity and the human becoming: principles of neuroarchaeology. \textit{Journal of Anthropological Sciences}, 88(4), pp.49–72.
%
%Malafouris, L., 2013. \textit{How Things Shape the Mind: A~Theory of Material Engagement}. Cambridge; London: The MIT Press.
%
%Malafouris, L., 2015. Metaplasticity and the Primacy of Material Engagement. \textit{Time and Mind}, [online] 8(4), pp.351–371. https://doi.org/10.1080/1751696X.2015.1111564.
%
%Mead, G.H., 1972. \textit{Mind, Self and Society: From the Standpoint of a~Social Behaviorist [1934}]. Chicago; London: University of Chicago Press.
%
%Medrano, M. and Urton, G., 2018. Toward the Decipherment of a~Set of Mid-Colonial Khipus from the Santa Valley, Coastal Peru. \textit{Ethnohistory}, [online] 65(1), pp.1–23. https://doi.org/10.1215/00141801-4260638.
%
%Merleau-Ponty, M., 2005. \textit{Phenomenology of Perception: An Introduction [1945}]. Routledge classics. Translated by C. Smith London; New York: Routledge.
%
%Newell, A., Simon, H.A. and Shaw, J.C., 1958. Elements of a~theory of human problem solving. \textit{Psychological Review}, 65(3), pp.151–166. https://doi.org/10.1037/h0048495.
%
%Orlikowski, W.J., 2007. Sociomaterial Practices: Exploring Technology at Work. \textit{Organization Studies}, [online] 28(9), pp.1435–1448. https://doi.org/10.1177/0170840607081138.
%
%Seaver, N., 2017. Algorithms as culture: Some tactics for the ethnography of algorithmic systems. \textit{Big Data \& Society}, [online] 4(2), pp.1–12. https://doi.org/10.1177/2053951717738104.
%
%Tomasello, M., 2014. \textit{A~Natural History of Human Thinking}: Cambridge, MA: Harvard University Press.
%
%Toscano, J., forthcoming. Intentionalities of the algorithm: historical practices and socio-computing infrastructures. A~philosophical account. \textit{Lo Sguardo}, 34(1).
%
%Vygotsky, L.S., 1978. Tool and Symbol in Child Development. In: M. Cole, V. John-Steiner, S. Scribner and E. Souberman, eds. \textit{Mind in Society: The Development of Higher Psychological Processes}. Cambridge: Harvard University Press. pp.19–30.
%\end{document}

>>>>

\section*{Initial Definitional Attempts}
\lettrine[loversize=0.13,lines=2,lraise=-0.03,nindent=0em,findent=0.2pt]%
{O}{}versimplified definitions of an algorithm are currently available and frequently used, but an algorithm is neither a~recipe nor a~rigidly constrained and procedural formulation. Limited conceptions that represent them as a~sort of entity or thing, a~series of steps that need to be applied, or a~simple technique that homogenizes a~process, lead to weak understandings of the deeper processes, transactions and dynamics that are at stake. Indeed, an algorithm can be a~problem-solving device, and this feature in itself can become a~point of entry to a~more complex analysis. After all, for engineers and computer scientists, ``an algorithm is an abstract, formalized description of a~computational procedure''
%\label{ref:RNDhginqouzXT}(Dourish, 2016, p.3).
\parencite[][p.3]{dourish_algorithms_2016}. %
 But even if sleek and apparently elegant, the problem with this definitional reduction is twofold.

@@
